---
title: "Pedro's Research"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

I'm Pedro Bergamo

## Links 

* [RMarkdown Intro](R/Rmarkdown_intro.html)

## Packages and data
```{r message=FALSE}
  library(leaflet)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(DT)
  library(scales) # install.packages("scales")
  library(patchwork)
```


```{r}
data_url <- "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e"

esc <- tryCatch(
    read.csv("data/escapement.csv"),
    error=function(cond) {
        message(paste("Escapement file does not seem to exist, so get it from the KNB."))
        esc <- read.csv(url(data_url, method = "libcurl"))
        return(esc)
    }
)

head(esc)
```


## Training dplyr and tidyr
```{r}

esc_separated <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), "-") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(SASAP.Region, Year, Species) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Sockeye", "Chinook", "Chum", "Pink", "Coho"))

head(esc_separated)
```

## ggplot lesson
```{r}
ggplot(esc_separated, mapping = aes(Species, escapement, fill = SASAP.Region)) + 
  geom_col()
```
### combining ggplot with tidyverse functions
```{r}
kodiak <- esc_separated %>% 
  filter(SASAP.Region == "Kodiak")

my_theme <- theme_classic() + 
  theme(legend.position = "bottom")

p1 <- ggplot(kodiak, aes(x=Year, y = escapement, color=Species)) +
  geom_line() +
  geom_point() +
  labs(x = "Year", y ="Escapement (num fish)", title = "Kodiak Salmon Escapement") +
  my_theme

ggsave("figures/kodiak-esc.png")
```

### facet_wrap function to generate multiple plots easily
```{r}
ggplot(esc_separated, aes(x=Year, y = escapement, color=Species))  + 
  geom_line() +
  geom_point() +
  scale_y_continuous(labels=comma) +
  facet_wrap(~SASAP.Region) +
  labs(x = "Year", y ="Escapement (num fish)") +
  my_theme
```
### patchwork package enables nice visualization of multiple plots
```{r}
p2<- ggplot(esc_separated, mapping=aes(x = Species, y = escapement, fill = Year)) +
  geom_col() +
  my_theme

p3<- ggplot(esc_separated, mapping=aes(x = Species, y = escapement, fill = SASAP.Region)) +
  geom_col() +
  my_theme

(p2 | p3)/p1 ##patchwork package - nice visualization of multiple ggplot graphs
```
### datatable function from DT package to create interactive tables
```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

datatable(locations)
```

## leaflet package

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~Longitude, lat = ~Latitude)
```

```{r}
leaflet(locations) %>% 
  addWMSTiles("https://www.gebco.net/data_and_products/gebco_web_services/web_map_service/mapserv?",
              layers = 'GEBCO_LATEST',
              attribution = "Imagery reproduced from the GEBCO_2014 Grid, version 20150318, www.gebco.net") %>%
  addCircleMarkers(lng = ~Longitude,
                   lat = ~Latitude,
                   popup = ~ Location,
                   radius = 5,
                   # set fill properties
                   fillColor = "salmon",
                   fillOpacity = 1,
                   # set stroke properties
                   stroke = T,
                   weight = 0.5,
                   color = "white",
                   opacity = 1)
```

